### 2016-03-15 paper.tex
Add Wosnik edits to Materials and Methods section

The rotor was 1.075 m in diameter and 0.8067 m tall, with blade chords that
taper from 0.04 m at the tips to 0.067 m at the roots, or half-span, varying the
chord-to-radius ratio $c/R$ from 0.07 to 0.12 (0.1 average). The rotor was
therefore approximately at the threshold between high and low solidity
\cite{Strickland1981,Fiedler2009}, which presents a unique validation case not
yet seen in the literature, similar in concept to the VAWT 850 \cite{Mays1990}
with its tapered blades, but with a moderately high $c/R$ to more accurately
represent typical MHK rotors, which presents a challenge for numerical modeling.

For investigating the effects of support strut drag losses, a set of cylindrical
covers were designed to slip over the struts, which provided a significant
increase in strut drag. Endcaps were also fabricated to allow the high drag
strut cover configuration to be operated without blades. A drawing of the strut
covers is shown in Fig.~\ref{fig:covers}.

@@

\paragraph{Synchronization.} The three data acquisition instrumentation
subsystems---motion controller, NI DAQ (performance measurements), and Vectrino+
(wake velocity measurements)---started sampling at precisely the same time each
tow, after being triggered by a digital signal edge from the motion controller.
This strategy retained synchronization for all performance signal samples (tow
speed, torque, drag, angular velocity), ensuring precise calculation of, e.g.,
power coefficient. Since there was also synchronization of the initial sample
from each of the three subsystems, correlation of events in the performance and
wake signals is also possible.

@@

Wake measurements were all performed at 1 m downstream, which corresponds to
$x/D = 0.93$. The cross-stream and vertical coordinates are shown in
Fig.~\ref{fig:coordinates}.

@@

For each tow speed, a relevant quasi-steady duration was selected by manually
inspecting a plot of the $C_P$ time series, and then truncated to include a
whole number of blade passages. Over this duration relevant statistics were
calculated.

Since the blockage ratio was relatively low (10\%), and a definitive blockage
correction for cross-flow turbines has not been established \cite{Cavagnaro2014}
blockage effects were not corrected for. Furthermore, most numerical modeling
efforts can easily incorporate finite domains, meaning blockage corrections
could complicate validation efforts.

@@

\begin{equation}
    C_P = \frac{T \omega}{\frac{1}{2} \rho A U_\infty^3}
\end{equation}
and
\begin{equation}
    C_D = \frac{F_\mathrm{drag}}{\frac{1}{2} \rho A U_\infty^2},
\end{equation}
where $\rho$ is the fluid density (1000 kg/m\textsuperscript{3}) and $A$ is
the turbine frontal area $DH$.

>>>>

The rotor was 1.075 m in diameter and 0.8067 m tall, with blade chords that
taper from 0.067 m at the roots, or half-span, to 0.04 m at the tips, varying
the chord-to-radius ratio $c/R$ from 0.12 to 0.07 (0.1 average). The rotor was
therefore approximately at the threshold between high and low solidity
\cite{Strickland1981,Fiedler2009}, which presents a unique validation case not
yet seen in the literature. The RM2 is conceptually similar to the VAWT 850
\cite{Mays1990} which also had tapered blades, but the RM2 has a moderately high
$c/R$ to more accurately represent typical MHK rotors, which presents a
challenge for numerical modeling.

For investigating the effects of support strut drag losses, a set of cylindrical
covers were designed to slip over the struts, which provided a deliberate and
significant increase in strut drag. Endcaps were also fabricated to allow the
high drag strut cover configuration to be operated without blades. A drawing of
the strut covers is shown in Fig.~\ref{fig:covers}.

@@

\paragraph{Synchronization.} The three data acquisition instrumentation
subsystems---motion controller, NI DAQ (performance measurements), and Vectrino+
(wake velocity measurements)---started sampling at precisely the same time each
tow, after being triggered by a digital signal edge sent from the motion
controller. This strategy retained synchronization for all performance signal
samples (tow speed, torque, drag, angular velocity), ensuring precise
calculation of, e.g., power coefficient. Since there was also synchronization of
the initial sample from each of the three subsystems, correlation of events in
the performance and wake signals is also possible.

@@

Wake measurements were all performed at 1 m downstream, which corresponds to
$x/D = 0.93$. The cross-stream and vertical coordinates are shown in
Fig.~\ref{fig:coordinates}. Altogether 750 tows were performed and included in
the experimental database.

@@

For each tow speed, a relevant quasi-steady duration was selected by manually
inspecting a plot of the $C_P$ time series. This interval was then truncated to
include a whole number of blade passages. Relevant statistics were then
calculated over this duration.

Since the blockage ratio was relatively low (10\%), and a definitive blockage
correction for cross-flow turbines has not been established \cite{Cavagnaro2014}
blockage effects were not corrected for. Furthermore, most numerical modeling
efforts can easily incorporate finite domains. Therefore, applying blockage
corrections to experimental data could actually complicate validation efforts.

@@

\begin{equation}
    C_P = \frac{T \omega}{\frac{1}{2} \rho A U_\infty^3}
\end{equation}
and
\begin{equation}
    C_D = \frac{F_\mathrm{drag}}{\frac{1}{2} \rho A U_\infty^2},
\end{equation}
where $\rho$ is the fluid density (assumed to be a nominal 1000
kg/m\textsuperscript{3}) and $A$ is the turbine frontal area $DH$.
